The effects of binarity on planet occurrence rates measured by transit surveys

### 2018-02-04 paper/main.tex
cite Hirsch 2017

This paper is organized as follows.  The next section enumerates the
various errors that arise from unrecognized binaries.  Then, in
Section~\ref{sec:simplest}, we develop an idealized model of a transit
survey in which all planets have identical properties, and all stars
are identical except that some fraction are in binary systems.  This
simple model motivates the derivation of a general formula, given in
Section~\ref{sec:general_formula}, that allows for more realistic
stellar and planetary populations.  We use this formula in
Section~\ref{sec:more_complicated} to explore more complicated but
nevertheless still analytic models.  In Section~\ref{sec:discussion}
we discuss the errors due to unrecognized binaries for specific cases
of current interest: the occurrence of Earth-like planets; the
apparent discrepancy between hot Jupiter occurrence rates in different
surveys; and the shape of the ``evaporation valley'' in the planet
radius distribution that was brought to light by
\citet{fulton_california-_2017}.  We review all the results in
Section~\ref{sec:conclusion}.

@@

Two other problems that may arise are not represented in
Equation~\ref{eq:occ_rate_simple}.  One is astrophysical: the true
occurrence rate of the desired type of planet may depend on whether
the host is a single star, the primary star of a binary, or the
secondary star of a binary.  Such differences could be caused by the
requirement for long-term dynamical stability, or differences in the
planet formation process.  When the search sample includes both
singles and binaries, the detected planets are thereby drawn from
different occurrence distributions~\citep[see, {\it
e.g.},][]{wang_occurrence_2015}.

@@

We can now understand that the observers are unwittingly reporting an
{\it apparent} occurrence rate $\Lambda\a(r\a)$, a function of
apparent radii.  The apparent occurrence rate is the same as the {\it
true} occurrence rate for transit signals of amplitude $\delta_{\rm
obs} = (r\a/R\a)^2$, but this latter quantity is typically not what
observers are interested in studying.

@@

The hypothesis that binarity might matter is grounded in the fact that
radial velocity and transit surveys treat binarity differently.
Radial velocity surveys typically reject both visual and spectroscopic
binaries \citep[\textit{e.g.},][]{wright_frequency_2012}.  Transit
surveys observe binaries, but the question of whether they were
searchable to begin with is left for later interpretation.  In
spectroscopic follow-up of transiting planet candidates, the
prevalence of astrophysical false-positives may also lead to a
tendency against confirming transiting planets in binary systems.  A
separate observational bias is that the RV surveys typically observe
chromospherically ``quiet'' stars, which may bias their detection
probability high~\citep{bastien_radial_2014}.

@@

\paragraph{Connecting high resolution imaging to occurrence rates}
One reason to address stellar multiplicity in transit surveys is that
it changes the interpretation of planet candidates on a
system-by-system level.  Consequently, high resolution imaging
campaigns have measured the multiplicities of almost all {\it Kepler}\
Objects of Interest~\citep{
  howell_speckle_2011,adams_adaptive_2012,adams_adaptive_2013,horch_observations_2012,
  horch_most_2014,lillo-box_multiplicity_2012,lillo-box_high-resolution_2014,dressing_adaptive_2014,
  law_robotic_2014,cartier_revision_2015,everett_high-resolution_2015,gilliland_hubble_2015,
  wang_influence_2015,wang_influence_2015-1,baranec_robo-ao_2016,ziegler_robo-ao_2017}.
The results of these programs have been collected
by~\citet{furlan_kepler_2017}, and they represent an important advance
in understanding the KOI sample's multiplicity statistics.  In
particular, they can be immediately applied to rectify binarity's
effects on the mass-radius diagram~\citep{furlan_densities_2017}.

These high resolution imaging campaigns are beginning to connect with
occurrence rate calculations: using~\citet{furlan_kepler_2017}'s
catalog, recent studies tested the effects of removing KOI hosts with
known companions, reducing contamination in the ``numerator'' of the
occurrence rate (\citealt{fulton_california-_2017};
\citealt{petigura_CKS_2017}).  While this is a good first step, the
denominator remains uncorrected.  To count the number of stars that
were searched, further work on understanding the multiplicity
statistics of KOI and non-KOI stars is necessary. It may be possible
to perform through a non-KOI star high-resolution imaging effort, or
perhaps by combining {\it Gaia} parallaxes with measured photometric
fluxes to identify binary systems.

\paragraph{Does a detected planet orbit the primary or secondary?}
A separate important step towards rectifying binarity's effects is
identifying the transit host star for detected planets in binaries.
The current wisdom is that, unless the planet can be confirmed to
orbit either component through analysis of the transit durations or
centroids, nothing substantive can be said about the likelihood of a
planet orbiting the primary vs. the
secondary~\citep[\textit{e.g.},][]{ciardi_understanding_2015,ziegler_robo-ao_2017}.

@@

\acknowledgements{
It was a pleasure sharing discussions about this study with T.
Barclay, W.  Bhatti, and F. Dai.  This work made use of NASA's
Astrophysics Data System Bibliographic Services.
%
\newline
%
\software{\texttt{numpy}~\citep{walt_numpy_2011}, 
\texttt{scipy}~\citep{jones_scipy_2001}, 
\texttt{matplotlib}~\citep{hunter_matplotlib_2007}, 
\texttt{pandas}~\citep{mckinney-proc-scipy-2010}
}
}

>>>>

This paper is organized as follows.  The next section enumerates the
various errors that arise from unrecognized binaries.  Then in
Section~\ref{sec:simplest}, we develop an idealized model of a transit
survey in which all planets have identical properties, and all stars
are identical except that some fraction are in binary systems.  This
simple model motivates the derivation of a general formula, given in
Section~\ref{sec:general_formula}, that allows for more realistic
stellar and planetary populations.  We use this formula in
Section~\ref{sec:more_complicated} to explore more complicated but
nevertheless still analytic models.  In Section~\ref{sec:discussion}
we discuss the errors due to unrecognized binaries for specific cases
of current interest: the occurrence of Earth-like planets; the
apparent discrepancy between hot Jupiter occurrence rates in different
surveys; and the shape of the ``evaporation valley'' in the planet
radius distribution that was brought to light by
\citet{fulton_california-_2017}.  We review all the results in
Section~\ref{sec:conclusion}.

@@

Two other problems that may arise are not represented in
Equation~\ref{eq:occ_rate_simple}.  One is astrophysical: the true
occurrence rate of the desired type of planet may depend on whether
the host is a single star, the primary star of a binary, or the
secondary star of a binary.  Such differences could be caused by the
requirement for long-term dynamical stability, or differences in the
planet formation process.  When the search sample includes both
singles and binaries, the detected planets are thereby drawn from
different occurrence 
distributions~\citep[{\it e.g.},][]{
    wang_occurrence_2015,kraus_impact_2016}.

@@

We can now understand that the observers are unwittingly reporting an
{\it apparent} occurrence rate $\Lambda\a(r\a)$, a function of
apparent radii.  The apparent occurrence rate is the same as the {\it
true} occurrence rate for transit signals of amplitude $\delta_{\rm
obs} = (r\a/R\a)^2$, but this latter quantity is less astrophysically
significant.

@@

Binarity might affect the inferred occurrence rates because 
radial velocity and transit surveys treat binaries differently.
Radial velocity surveys typically reject both visual and spectroscopic
binaries \citep[\textit{e.g.},][]{wright_frequency_2012}.  Transit
surveys observe binaries, but the question of whether they were
searchable to begin with is left for later interpretation.  In
spectroscopic follow-up of transiting planet candidates, the
prevalence of astrophysical false-positives may also lead to a
tendency against confirming transiting planets in binary systems.  A
separate observational bias is that the RV surveys typically observe
chromospherically ``quiet'' stars, which may bias their detection
probability high~\citep{bastien_radial_2014}.

@@

\paragraph{Connecting high resolution imaging to occurrence rates}
One reason to address stellar multiplicity in transit surveys is that
it can radically alter the interpretation of planet candidates on a
system-by-system level.  Consequently, high resolution imaging
campaigns have measured the multiplicities of almost all {\it Kepler}\
Objects of Interest~\citep{
  howell_speckle_2011,adams_adaptive_2012,adams_adaptive_2013,horch_observations_2012,
  horch_most_2014,lillo-box_multiplicity_2012,lillo-box_high-resolution_2014,dressing_adaptive_2014,
  law_robotic_2014,cartier_revision_2015,everett_high-resolution_2015,gilliland_hubble_2015,
  wang_influence_2015,wang_influence_2015-1,baranec_robo-ao_2016,ziegler_robo-ao_2017}.
The results of these programs have been collected
by~\citet{furlan_kepler_2017}, and they represent an important advance
in understanding the KOI sample's multiplicity statistics.  
An immediate application is to reassess the radii of detected planets.
Doing so, \citet{hirsch_assessing_2017} find that for their sample of KOIs
in binaries, the planet radii are underestimated by factors $r/r\a = 1.17$
if all planets orbit primaries, and $r/r\a = 1.65$ if detected
planets are equally likely to orbit primaries and singles.
%In particular, they can be immediately applied to rectify binarity's
%effects on the mass-radius diagram~\citep{furlan_densities_2017}.

These results are beginning to connect with
occurrence rate calculations: using~\citet{furlan_kepler_2017}'s
catalog, recent studies have tested the effects of removing KOI hosts with
known companions, reducing contamination in the ``numerator'' of the
occurrence rate (\citealt{fulton_california-_2017};
\citealt{petigura_CKS_2017}).  
While these are good first steps, the
denominator remains uncorrected.  To count the number of stars that
were searched, further work on understanding the multiplicity
statistics of KOI and non-KOI stars is necessary. In particular,
a non-KOI star high-resolution imaging effort could help estimate
the systematic errors incurred by assuming all surveyed stars are
single.
An alternate approach would be to combine {\it Gaia}
parallaxes with measured photometric fluxes and identify binary systems
as those that are 
overluminous~\citep[\textit{e.g.},][]{widmark_inferring_2018}.

\paragraph{Does a detected planet orbit the primary or secondary?}
An important step towards rectifying binarity's effects is
identifying the transit host star for detected planets in binaries.
The current wisdom is that, unless the planet can be confirmed to
orbit either component through analysis of the transit durations or
centroids, little can be said about the likelihood of a
planet orbiting the primary vs. the
secondary~\citep[\textit{e.g.},][but see~\citealt{hirsch_assessing_2017}'s 
approach to weighting radius corrections by planet 
occurrence]{ciardi_understanding_2015,ziegler_robo-ao_2017}.

@@

\acknowledgements{
It was a pleasure discussing this study with T.
Barclay, W.  Bhatti, and F. Dai.  This work made use of NASA's
Astrophysics Data System Bibliographic Services.
%
\newline
%
\software{\texttt{numpy}~\citep{walt_numpy_2011}, 
\texttt{scipy}~\citep{jones_scipy_2001}, 
\texttt{matplotlib}~\citep{hunter_matplotlib_2007}, 
\texttt{pandas}~\citep{mckinney-proc-scipy-2010}
}
}
